Statistics Project

### 2022-12-10 rough_analysis.Rmd
Dialouge for document + Additional model components for A2

For the first dataset, we are looking at houses are are only within the [NAmes, Edwards, BrkSide neighborhoods] and are looking at how

SalePrice ~ SqFt (GrLIvArea) + Neighborhood

and if SalesPrice depends on neighborhood.

```{r echo=FALSE}
library(tidyverse)
library(plotly)
library(car)

housing = read.csv(file = "train.csv")
numeric_housing = housing %>% select_if(is.numeric)
a1_set = housing %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
```

Initializing any required packages and calling in the data with some light filtering, we can take a look at the dataset and visualize what the datalooks like with some cursory visuals.

@@

We can see in the dataset that there is a difference in slope/intercept values per group

```{r}
#Comparison of Means/Coefficients to signify a difference in sample groups from different neighborhoods.
n_means <- aov(SalePrice ~ Neighborhood, data = a1_set)
mean_diff <- TukeyHSD(n_means, conf.level = 0.95)
mean_diff
fit <- lm(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
```

@@

@@

library(caret)
train.control = trainControl(method = "cv", number = 10)
step.model = train(SalePrice ~.-GarageArea, data = housing %>% select_if(is.numeric) %>% drop_na(), method = "leapBackward", trControl = train.control)
step.model$modelInfo
```

For the process of forward feature selection or backwards feature elimination, we can use the olsrr package to do most of the calculations for us. This can be seen as follows:

```{r}
#install.packages("olsrr")
library(olsrr)

@@

fit_numeric = lm(SalePrice ~., data = housing %>% select_if(is.numeric))

@@

```{r}
for(i in housing %>% names()){
  print(nlevels(housing %>% select(i)))
}

lapply(housing[character_names], nlevels)
```

>>>>

For the first analysis, we are looking at houses are are only within the locale of [NAmes, Edwards, BrkSide neighborhoods] and at how Sale Price can possibly change
with respect to the Size of the living area and the Neighborhood in which the property is located, which can be visualized with the equation:

SalePrice ~ SqFt (GrLIvArea) + Neighborhood + GrLIvArea:Neighborhood

Pulling the data into R and performing some rudimentary cleaning;

```{r include=FALSE}
library(tidyverse)
library(plotly)
library(car)
library(caret)
library(olsrr)
```

```{r RMSE function}
ModelRMSE <- function(res){
  retRMSE <- sqrt(mean(res^2))
  return(retRMSE)
}
```

```{r CVPress function}
CVPress <- function(input_model){
  pred_res = resid(input_model)/(1 - lm.influence(input_model)$hat)
  press <- sum(pred_res^2)
  return(press)
}
```

```{r}
housing = read.csv(file = "train.csv")
test_housing = read.csv(file = "test.csv")

numeric_housing = housing %>% select_if(is.numeric)
a1_set = housing %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
```

We can take a look at the dataset and visualize what the datalooks like with some cursory visuals. During this step, we want to analyze any heavy correlation that the dataset might have, and account for it in any of our developing models;

@@

From visual inspection, we can identify that there is a visual difference in slope/intercept values per group, and can use methods to identify and quantify what these values are.

```{r}
#Comparison of Means/Coefficients to signify a difference in sample groups from different neighborhoods.
### comparison of means for the groups
n_means <- aov(SalePrice ~ Neighborhood, data = a1_set)
mean_diff <- TukeyHSD(n_means, conf.level = 0.95)
mean_diff

### linear model with interactions explained per group
fit <- lm(SalePrice ~ log(GrLivArea) + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
summary(fit)

#RMSE: 
ModelRMSE(fit$residuals)
```

@@

For Analysis 2, we need to generate 4 models using variations of parameters selected by Forwards Parameter Selection, Backwards Elimination, Stepwise Selection, and a 
personalized custom approach.

Using these paramters, we can generate Adjusted R^2 values, CV Press statistics, and finally submit our model to the Kaggle website to obtain a "Kaggle Score" for each model.

This process begins as follows:

@@

```{r}
#bad approach, maybe, I can't get it to work 
#train.control = trainControl(method = "cv", number = 10)
#step.model = train(SalePrice ~.-GarageArea, data = housing %>% select_if(is.numeric) %>% drop_na(), method = "leapBackward", trControl = train.control)
#step.model$modelInfo
```

For the process of forward feature selection or backwards feature elimination, we can use the `olsrr` package to do most of the selection work for us. This can be seen as follows:

```{r}
#install.packages("olsrr")

@@

fit_numeric = lm(SalePrice ~., data = housing %>% select_if(is.numeric)) # does not account for categorical parameters !!!

@@

From these methods we can pull our the corresponding models and analyze the statistics as well as the parameters that we chosen, to verify that first, there was a reduction in selection in the parameters for the LM and second,

This step might not be necessary but could improve the quality of the model. Analysis shows a marginal improvement to R^2 and marginal deteriorate to PRESS Stat when rm'ing outliers.

Outlier ID# : 692, 1183,  Excessively Expensive > 700k
Outlier ID# : 314, 336, 256, 707    Excessively Large > 100k sqr ft.

```{r}
#I'm not sure if this is the best way to do it, but it highlights the requirement of the transform

boxplot(numeric_housing) #<-visualize outliers

#filtered_nh = numeric_housing[-c(692, 1183, 314, 336, 256, 707), ] #filters out outliers 
#filtered_nh$SalePrice = log(filtered_nh$SalePrice) 
#filtered_nh$LotArea = log(filtered_nh$LotArea)

transformed_numeric = numeric_housing
transformed_numeric$SalePrice = log(transformed_numeric$SalePrice)
transformed_numeric$LotArea = log(transformed_numeric$LotArea)

boxplot(transformed_numeric)

transformed_fit = lm(SalePrice ~., data = transformed_numeric)

transForward = ols_step_forward_p(transformed_fit, penter = 0.05, details = FALSE)

ModelRMSE(transForward$model$residuals) #see defined functions at top of document
CVPress(transForward$model)

#boxplot(filtered_nh)

#filtered_fit_numeric = lm(SalePrice ~., data = filtered_nh)                               #outlier filtered approach

#f_forward = ols_step_forward_p(filtered_fit_numeric, penter = 0.05, details = FALSE)

#ModelRMSE(f_forward$model$residuals) #see defined functions at top of document
#CVPress(f_forward$model)
```
